Hierarchically modelling \emph{Kepler} dwarfs and subgiants to improve inference of stellar properties with asteroseismology

### 2020-10-27 paper/main.tex
WJC - introduction chapter 2 grammar
Added comment

Typically, stars are modelled on a star-by-star basis with estimates of stellar properties including some assumptions handled by simple scaling relations and solar calibrations. % assumptions made about the bulk physical quantities based on empirical relations or solar calibrations. 
For instance, a helium ($Y$) to heavy-element ($Z$) ratio ($\Delta Y / \Delta Z$) and mixing-length theory parameter ($\mlt$) are often assumed. However, there has been little effort to account for the population distribution of such quantities. Assuming $Y$ and $\mlt$ can result in inaccurate inference and systematics \citep{Valle.DellOmodarme.ea2015}. Independently modelling $Y$ and $\mlt$ can also be computationally demanding and requires high-precision observations in order to return meaningful stellar properties.

@@

Using five independent pipelines, \citetalias{Serenelli.Johnson.ea2017} determined values for global asteroseismic parameters -- the large frequency separation $\dnu$ and the frequency at maximum power, $\numax$ with median uncertainties of 1.7 per cent and 4 per cent respectively. We chose to adopt the $\dnu$ determined in their work as inputs for our method. They also used $\metallicity$ published in Data Release 13 \citep[DR13;][]{Albareti.AllendePrieto.ea2017} of the APOGEE stellar abundances pipeline \citep[ASPCAP;][]{GarciaPerez.AllendePrieto.ea2016} with an additional uncertainty of \SI{0.1}{\dex}. For their preferred set of results, they adopted a $\teff$ scale from the Sloan Digital Sky Survey (SDSS) \emph{griz}-band photometry \citep{Pinsonneault.An.ea2012} with a median uncertainty of \SI{70}{\kelvin}.

We removed more evolved stars from the APOKASC sample by cutting those with $\log g < \SI{3.8}{\dex}$. We then kept stars within 1-$\sigma$ of $-0.5 < \metallicity < \SI{0.5}{\dex}$ to remove metal-poor stars. Stars with $M \gtrsim \SI{1.2}{\solarmass}$ are understood to have a convective, hydrogen-burning core, with some dependence on the choice of stellar physics \citep{Appourchaux.Antia.ea2015}. Stellar models with a convective core require the treatment of extra stellar physics such as overshooting, which is beyond the scope of this work. Therefore, we keep only stars with masses within 1-$\sigma$ of \SIrange{0.8}{1.2}{\solarmass} from the preferred set of results of \citetalias{Serenelli.Johnson.ea2017}.

@@

Our principle goal was to imporve inference of fundamental stellar parameters for our sample of low-mass stars. To achieve this, we constructed a hierarchical Bayesian model (HBM) which utilises a prior assumption of the population distribution to share information between the stars. Based on the work of Davies et al. (in prep.), our HBM is a generative model which requires a function to map bulk stellar parameters to their observables.

@@

\begin{table}
	\centering
	\caption{Stellar model grid parameters for training and test datasets.}
	\label{tab:grid}
	\begin{tabular}{cccc} % four columns, alignment for each
		\toprule
		\multicolumn{4}{c}{\textbf{Stellar model grid}}\\
		\midrule
		Input Parameter & Range & Increment & $N_{\rm track}$\\
        \midrule
	$M$ ($M_{\odot}$)  & 0.80 -- 1.20 &  0.01& \num{41}\\
        $\rm{[M/H]}$ (dex) & -0.5 -- 0.2/0.25 -- 0.5 & 0.1/0.05 & \num{14}\\
        	$Y_{\rm init}$ & 0.24 -- 0.32 & 0.02 & \num{5}\\
        $\alpha_{\rm{mlt}}$  & 1.5 -- 2.5&  0.2 & \num{6}\\
        \midrule
        \textbf{Total} & & & \num{17220}\\
        \bottomrule
        %Other physics & \multicolumn{2}{Scheme}\\
        %\hline
        %Diffusion & \multicolumn{2}{Yes}\\
        %Overshooting & \multicolumn{2}{N/A}\\
        %\hline
%     \multicolumn{3}{p{.4\textwidth}}{$^{\rm a}Y_0$ = 0.249 and $\frac{\Delta Y}{\Delta Z}$ = 1.33 was adopted.}  
	\end{tabular}
\end{table}

@@

Theoretical stellar oscillations were calculated with the \textsc{GYRE} code (version 5.1), which was developed by \citet{Townsend.Teitler2013}. We computed radial modes (for $\ell$ = 0) by solving the adiabatic stellar pulsation equations with the structural models generated by \textsc{MESA}. We determined the asteriseismic large separation ($\dnu$) for each model with theoretical radial modes to avoid the systematic offset of the scaling relation. We derived $\Delta \nu$ with the approach given by \citet{White.Bedding.ea2011}, which is a weighted least-squares fit to the radial frequencies as a function of $n$. We chose to ignore the well known, yet poorly characterised impact of modelled oscillation mode inaccuracies in the near-surface region of the star \citep{Kjeldsen.Bedding.ea2008, Ball.Gizon2014, Sonoi.Samadi.ea2015}. % This presents only a small effect when considering the average large frequency spacing, $\dnu$ and is beyond the scope of this paper.

@@

Using notation which represents the distribution of some parameter $x \sim g$ as equivalent to $p(x) \propto g(x)$ where $g(x)$ is some non-normalised probability density function, we write the priors for $\boldsymbol{\theta}_i$ as,
%
\begin{align*}
    f_{\mathrm{evol}, i} &\sim 0.01 + 1.99 \cdot \mathcal{B}(1.2, 1.2),\\
    M_i &\sim 0.8 + 0.4 \cdot \mathcal{B}(1.2, 1.2),\\
    \alpha_{\mathrm{MLT}, i} &\sim 1.5 + \mathcal{B}(1.2, 1.2),\\
    Y_{\mathrm{init}, i} &\sim 0.24 + 0.08 \cdot \mathcal{B}(1.2, 1.2),\\
    Z_{\mathrm{init}, i} &\sim 0.005 + 0.035 \cdot \mathcal{B}(1.2, 1.2),\\
\end{align*}
%
where each beta distribution is scaled to cover the boundaries of the grid of stellar models computed in Section \ref{sec:grid}.

@@

We built another hierarchical model similar to the PP model except that $\mlt$ is max-pooled. In other words, we assumed that the mixing length must be the same value for every star in the sample, but still allowed it to freely vary. Thus the hyperparameters are now, $\boldsymbol{\phi} = \{\Delta Y/\Delta Z, Y_P, \sigma_Y, \mlt\}$. The posterior distribution of the model takes the same form as in Equation \ref{eq:hbmbayes} except that the mixing-length theory parameter for the $i$-th star is,
%
\begin{equation}
    \alpha_{\mathrm{MLT}, i} = \mlt,
\end{equation}
%
where,
\begin{equation}
    \mlt \sim 1.5 + \mathcal{B}(1.2, 1.2).
\end{equation}
% %
% \begin{equation}
%     p(\boldsymbol{\alpha}_\mathrm{MLT} | \mlt) = \prod_{i=1}^{N_\mathrm{stars}} \delta (\alpha_{\mathrm{MLT}, i} | \mlt)
% \end{equation}
% %
% where $\delta(x | \alpha)$ is defined as,
% %
% \begin{equation}
%     \delta(x | \alpha) = \begin{cases}
%         + \infty, &\quad x = \alpha\\
%         0, &\quad x \neq \alpha
%     \end{cases}
% \end{equation}
In other words, $\mlt$ is a free parameter in the model but is assumed to be the same in all stars.

@@

For both the PP and MP models, we ran versions with and without data for the Sun included in the APOKASC population, referred to as PPS and MPS respectively. We adopted standard solar data as shown in Table \ref{tab:sun} with uncertainties conservatively limited to the accuracy of the neural network. We also adopted $\dnu=\SI{135.1(2)}{\micro\hertz}$ from \citet{Huber.Bedding.ea2011} with a standard deviation representative of variations in measurements of the solar $\dnu$ \citep{Broomhall.Chaplin.ea2011}.

@@

\begin{table}
    \centering
    \caption{Solar input data.}
    \label{tab:sun}
    \input{sun_inputs.tex}
\end{table}

@@

In Table \ref{tab:np}, we preset results for the remaining 65 stars from the NP model. Running the NP model with synthetic stars resulted in unreliable uncertainties (see Appendix \ref{sec:test-stars}). This was because the boundary of the priors in $Y_\mathrm{init}$ and $\mlt$ truncated the posterior distribution leading to underestimated uncertainties and skewing their posterior means towards the centre of their priors. Therefore, we present the NP results only for comparison purposes, but we exclude it from further discussion.

>>>>

Typically, stars are modelled on a star-by-star basis with estimates of stellar properties including some assumptions handled by simple scaling relations and solar calibrations. % assumptions made about the bulk physical quantities based on empirical relations or solar calibrations. 
For instance, a helium ($Y$) to heavy-element ($Z$) ratio ($\Delta Y / \Delta Z$) and mixing-length theory parameter ($\mlt$) are often assumed. However, there has been little effort to account for the population distribution of such quantities. Assuming values for $Y$ and $\mlt$ can result in inaccurate inference and systematics on the inferred stellar parameters \citep{Valle.DellOmodarme.ea2015}. Independently modelling $Y$ and $\mlt$ can also be computationally demanding and requires high-precision observations in order to return meaningful stellar properties.

@@

Using five independent pipelines, \citetalias{Serenelli.Johnson.ea2017} determined values for global asteroseismic parameters -- the large frequency separation $\dnu$ and the frequency at maximum power, $\numax$ with median uncertainties of 1.7 per cent and 4 per cent respectively. We chose to adopt the $\dnu$ determined in their work as inputs for our method. They also used $\metallicity$ published in Data Release 13 \citep[DR13;][]{Albareti.AllendePrieto.ea2017} of the APOGEE stellar abundances pipeline \citep[ASPCAP;][]{GarciaPerez.AllendePrieto.ea2016} with uncertainties of \SI{0.1}{\dex}. For their preferred set of results, they adopted a $\teff$ scale from the Sloan Digital Sky Survey (SDSS) \emph{griz}-band photometry \citep{Pinsonneault.An.ea2012} with a median uncertainty of \SI{70}{\kelvin}.

% WJC - cut may exclude more metal-rich stars?
We removed more evolved stars from the APOKASC sample by cutting those with $\log g < \SI{3.8}{\dex}$. We then kept stars within 1-$\sigma$ of $-0.5 < \metallicity < \SI{0.5}{\dex}$ to remove metal-poor stars. Main sequence stars with $M \gtrsim \SI{1.2}{\solarmass}$ are understood to have a convective, hydrogen-burning core, with some dependence on the chemical composition and choice of stellar physics \citep{Appourchaux.Antia.ea2015}. Stellar models with a convective core require the treatment of extra stellar physics such as overshooting, which is beyond the scope of this work. Therefore, we keep only stars with masses within 1-$\sigma$ of \SIrange{0.8}{1.2}{\solarmass} from the preferred set of results of \citetalias{Serenelli.Johnson.ea2017}.

@@

Our principle goal was to improve inference of fundamental stellar parameters for our sample of low-mass stars. To achieve this, we constructed a hierarchical Bayesian model (HBM) which utilises a prior assumption of the population distribution to share information between the stars. Based on the work of Davies et al. (in prep.), our HBM is a generative model which requires a function to map bulk stellar parameters to their observables.

@@

\begin{table}
	\centering
	\caption{Stellar model grid parameters for training and test datasets. $N_\mathrm{track}$ are the numbers of stellar evolutionary tracks computed for each dimension of the grid, multiplied to a total of \num{17220} tracks.}
	\label{tab:grid}
	\begin{tabular}{cccc} % four columns, alignment for each
		\toprule
		\multicolumn{4}{c}{\textbf{Stellar model grid}}\\
		\midrule
		Input Parameter & Range & Increment & $N_{\rm track}$\\
        \midrule
	$M$ ($M_{\odot}$)  & 0.80 -- 1.20 &  0.01& \num{41}\\
        $\rm{[M/H]}$ (dex) & -0.5 -- 0.2/0.25 -- 0.5 & 0.1/0.05 & \num{14}\\
        	$Y_{\rm init}$ & 0.24 -- 0.32 & 0.02 & \num{5}\\
        $\alpha_{\rm{mlt}}$  & 1.5 -- 2.5&  0.2 & \num{6}\\
        \midrule
        \textbf{Total} & & & \num{17220}\\
        \bottomrule
        %Other physics & \multicolumn{2}{Scheme}\\
        %\hline
        %Diffusion & \multicolumn{2}{Yes}\\
        %Overshooting & \multicolumn{2}{N/A}\\
        %\hline
%     \multicolumn{3}{p{.4\textwidth}}{$^{\rm a}Y_0$ = 0.249 and $\frac{\Delta Y}{\Delta Z}$ = 1.33 was adopted.}  
	\end{tabular}
\end{table}

@@

Theoretical stellar oscillations were calculated with the \textsc{GYRE} code (version 5.1), which was developed by \citet{Townsend.Teitler2013}. We computed radial modes (for $\ell$ = 0) by solving the adiabatic stellar pulsation equations with the structural models generated by \textsc{MESA}. We determined the asteroseismic large separation ($\dnu$) for each model with theoretical radial modes to avoid the systematic offset of the scaling relation. We derived $\Delta \nu$ with the approach given by \citet{White.Bedding.ea2011}, which is a weighted least-squares fit to the radial frequencies as a function of $n$.

We chose to ignore the well known, yet poorly characterised impact of modelled oscillation mode inaccuracies in the near-surface region of the star \citep{Kjeldsen.Bedding.ea2008, Ball.Gizon2014, Sonoi.Samadi.ea2015}. % This presents only a small effect when considering the average large frequency spacing, $\dnu$ and is beyond the scope of this paper.

@@

Using notation which represents the distribution of some parameter $x \sim q$ as equivalent to $p(x) \propto q(x)$ where $q(x)$ is some non-normalised probability density function, we write the priors for $\boldsymbol{\theta}_i$ as,
%
\begin{align*}
    f_{\mathrm{evol}, i} &\sim 0.01 + 1.99 \cdot \mathcal{B}(1.2, 1.2),\\
    M_i &\sim 0.8 + 0.4 \cdot \mathcal{B}(1.2, 1.2),\\
    \alpha_{\mathrm{MLT}, i} &\sim 1.5 + \mathcal{B}(1.2, 1.2),\\
    Y_{\mathrm{init}, i} &\sim 0.24 + 0.08 \cdot \mathcal{B}(1.2, 1.2),\\
    Z_{\mathrm{init}, i} &\sim 0.005 + 0.035 \cdot \mathcal{B}(1.2, 1.2),\\
\end{align*}
%
where each beta distribution is scaled to cover the boundaries of the grid of stellar models computed in Section \ref{sec:grid}.

@@

We built another hierarchical model similar to the PP model except that $\mlt$ is max-pooled. In other words, we assumed that the mixing length must be the same value for every star in the sample, but still allowed it to freely vary. Thus the hyperparameters are now, $\boldsymbol{\phi} = \{\Delta Y/\Delta Z, Y_P, \sigma_Y, \mlt\}$. The posterior distribution of the model takes the same form as in Equation \ref{eq:hbmbayes} except that the mixing-length theory parameter for the $i$-th star is,
%
\begin{equation}
    \alpha_{\mathrm{MLT}, i} = \mlt,
\end{equation}
%
where,
\begin{equation}
    \mlt \sim 1.5 + \mathcal{B}(1.2, 1.2).
\end{equation}
% %
% \begin{equation}
%     p(\boldsymbol{\alpha}_\mathrm{MLT} | \mlt) = \prod_{i=1}^{N_\mathrm{stars}} \delta (\alpha_{\mathrm{MLT}, i} | \mlt)
% \end{equation}
% %
% where $\delta(x | \alpha)$ is defined as,
% %
% \begin{equation}
%     \delta(x | \alpha) = \begin{cases}
%         + \infty, &\quad x = \alpha\\
%         0, &\quad x \neq \alpha
%     \end{cases}
% \end{equation}
chosen such that $\mlt$ is confined to the boundaries of the grid of stellar models ($1.5 < \mlt < 2.5$). In other words, $\mlt$ is a free parameter in the model but is assumed to be the same in all stars.

@@

For both the PP and MP models, we ran versions with and without data for the Sun included in the APOKASC population, referred to as PPS and MPS respectively. We adopted standard solar data as shown in Table \ref{tab:sun} with uncertainties conservatively limited to the accuracy of the ANN for $R$, $L$ and representative of variation in the literature for $\teff$. We also adopted $\dnu=\SI{135.1(2)}{\micro\hertz}$ from \citet{Huber.Bedding.ea2011} with a standard deviation representative of variations in measurements of the solar $\dnu$ \citep{Broomhall.Chaplin.ea2011}.

\begin{table}
    \centering
    \caption{Solar input data where $\mu$ and $\sigma$ are the means and standard deviations respectively.}
    \label{tab:sun}
    \input{sun_inputs.tex}
\end{table}

@@

@@

In Table \ref{tab:np}, we preset results for the remaining 65 stars from the NP model. Running the NP model with synthetic stars resulted in unreliable uncertainties (see Appendix \ref{sec:test-stars}). This was because the boundary of the priors in $Y_\mathrm{init}$ and $\mlt$ truncated the posterior distribution leading to underestimated uncertainties and skewing their posterior means towards the centre of their priors. Therefore, we present the NP results only for comparison purposes, but we exclude them from further discussion.
